---
title: " Data Insights"
format: html
editor: visual
---
## Glimpse of Episodes
```{r}
library(tidyverse)
library(ggplot2)
twd_episodes <- read.csv("the_walking_dead_episodes.csv")
twd_episodes$season <- as.double(twd_episodes$season)
glimpse(twd_episodes)
```

## Comparing Viewership Over Time 

```{r, eval=T, echo=F, error=T}
twd_episodes$original_air_date <- as.Date(twd_episodes$original_air_date, format = "%Y-%m-%d")
# extract year
twd_episodes$air_year<- format(twd_episodes$original_air_date, "%Y")
# plot
ggplot(data=twd_episodes, #data layer
       mapping = aes(x=air_year, y=us_viewers)) +
  geom_point() + #geometric layer
  labs(title="Viewers in the US over Time",
       x= "Air Date(year)", y= "Viewcount")
```

### Comparing Viewership Over Time, grouped by Season
```{r, eval=T,echo=F}
ggplot(data=twd_episodes, #data layer
       mapping = aes(x=air_year, y=us_viewers, color=factor(season))) +
  geom_point() + #geometric layer
  labs(title="Viewers in the US over Time",
       x= "Air Date(year)", y= "Viewcount")
```
## Who are the Directors and what is their impact to the show?
``` {r}
twd_episodes %>% group_by(directed_by) %>% summarise(episode_count=n(),average_viewcount=mean(us_viewers)) %>% arrange(desc(episode_count))
```

For More in-depth breakdowns of what each director does, scroll through here:
<iframe height="400" width="100%" frameborder="no" src="https://starrynad.shinyapps.io/TwdDirector/">
</iframe> 


### Seasonal change

## Who are the Writers and what is their impact? 
- count
- mean viewership

### Seasonal change

## Average IMDB score per Season

## Average IMDB score per show Writer 

## Average IMDB score per show Director 

## Other factors that affect popularity 

### Shortest and Longest Season 

# Conclusion 
- most popular season 
- most popular writer 
- most popular director 
- IMDB vs Viewership 
- Length of Show potential downfall